---
title: "R: practice writing scripts: for, if, user-defined funcions, ggplot2"
author: "Anna Williford"
date: "9/12/2019"
output: html_document
---
  
  
I was given `gapminder.txt` dataset and asked to do the following:

**For every unique country in gapminder.txt dataset, if the country starts with 'Ma', use ggplot2 package to plot life expectancy(y axis) over years(x axis).**  

Here are the steps I need to implement in R to complete this task.  
1. Load `ggplot2` library  
2. Load `gapminder.txt`  
3. Create `country_list` that contains country names from `gapminder.txt`. 
4. Start a `for` loop to iterate over every element of `country_list`
5. Use `if` statement to check if the country name starts with "Ma".    
6. If it does, subset gapminder dataset to select information about just that country.   
7. Make ggplot to plot life expectancy on y-axis and years on x-axis using subset of the data from step 6.  
8. Close `if`statement and `for` loop.


Here is a scaffold for a programm:
```
#1 load library
library(ggplot2)

#2 load data
gapminder<-read.table("gapminder.txt", header = TRUE)

#3 create list of countries to loop over
country_list=

#4 loop over country_list
for(country in country_list){
  #5 if(country starts with "Ma"){
     #5 subset gapminder
     #6 make ggplot
   }
}
```
Here is the implementation of these steps:  

```{r, plots}
#load packages
library(ggplot2)

#read data ito R
gapminder<-read.table("gapminder.txt", header = TRUE)

#make country_list to loop over
country_list=unique(gapminder$country)

#loop over counry_list
for(country in country_list) {
  if(startsWith(country,"Ma")){
    #subset data to get info about the country if its name starts with 'Ma'
    countryData<-gapminder[gapminder$country==country,]
    #make ggplot with countryData
    print(ggplot(data=countryData)+aes(x=year,y=lifeExp)+geom_line(color="red")
          +labs(title=country))
    
  }
}
```
If you want to save each plot to a separate file, make a unique name for every plot and save with `ggsave`

```{r, loop_if_plot, eval=FALSE}
#load packages
library(ggplot2)

#read data ito R
gapminder<-read.table("gapminder.txt", header = TRUE)

#make country_list to loop over
country_list=unique(gapminder$country)

#loop over counry_list
for(country in country_list) {
  if(startsWith(country,"Ma")){
    #subset data to get info about the country if its name starts with 'Ma'
    countryData<-gapminder[gapminder$country==country,]
    #make ggplot with countryData
    myplot<-ggplot(data=countryData)+aes(x=year,y=lifeExp)+geom_line(color="red")
    #create a unique file name to save plot to
    f_name=paste0(country,"_plot.png")
    #save plot
    ggsave(myplot, filename=f_name)
  }
}
```
**Note** The script would be easier to read if subsetting was done with a user-defined function:

```{r, subsetting function, eval=FALSE}
getCountryData<-function(countryName, df){
  countryData<-df[df$country==countryName,] 
  return(countryData)
}
```

Then, we can rewrite the above script like this:

```{r, loop_if_plot_withFunction, eval=FALSE}
#load packages
library(ggplot2)

#define functions
getCountryData<-function(countryName, df){
  countryData<-df[df$country==countryName,] 
  return(countryData)
}

#read data ito R
gapminder<-read.table("gapminder.txt", header = TRUE)

#make country_list to loop over
country_list=unique(gapminder$country)

#loop over counry_list
for(country in country_list) {
  if(startsWith(country,"Ma")){
    #subset data to get info about the country if its name starts with 'Ma'
    countryData<-getCountryData(country,gapminder)
    #make ggplot with countryData
    myplot<-ggplot(data=countryData)+aes(x=year,y=lifeExp)+geom_line(color="red")
    #create a unique file name to save plot to
    f_name=paste0(country,"_plot.png")
    #save plot
    ggsave(myplot, filename=f_name)
  }
}
```
